---
title: "Estimation of CCDPH's Seasonal Populations"
author: "C. Scott Smith, PhD AICP"
date: "3/26/2021"
output:
  word_document:
    reference_docx: StylesTemplate.docx
    keep_md: true
csl: chicago-author-date.csl
bibliography: references.bib
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(censusapi)
library(tigris) 
library(tidyverse)
library(tidyr)
library(dplyr) 
library(sf) 
library(clipr)
library(readxl)
library(leaflet)
library(scales)
library(ggmap)
library(data.table)
library(XML)
library(reshape2)
library(ggplot2)

# https://fishandwhistle.net/post/2020/getting-started-zotero-better-bibtex-rmarkdown/
# https://cran.r-project.org/web/packages/ymlthis/vignettes/yaml-fieldguide.html
# https://www.zotero.org/styles
# https://retorque.re/zotero-better-bibtex/installation/
# https://retorque.re/zotero-better-bibtex/install.packages('tinytex')
# install.packages('tinytex')
# tinytex::install_tinytex()

```

## Temporary populations and public health planning

As public health programs become more targeted toward specific
population segments, there is a need to better understand the behaviors
of temporary populations to increase intervention efficacy. Of
particular concern here are the longer-term mobility patterns of
seasonal residents, which have the potential to substantively alter the
size and composition of Cook County Department of Public Health's
(CCDPH) population at different times of the year. Accurate estimates of
the size, distribution, timing and characteristics of temporary
populations has been especially important for COVID-19 vaccination
planning and programs, which require accurate counts of the size and
locations of phase-specific priority groups. This article: (1) briefly
summarizes some challenges associated with estimating seasonal
populations; (2) proposes a methodology for counting seasonal
residents--specifically 'snowbird' populations--using American Community
Survey (ACS) data; and (3) applies the methodology to develop
place-level counts of seasonal populations within the CCDPH
jurisdiction.

## Challenges with estimating temporary populations

Temporarily occupied units are sample units occupied at the time of
interview entirely by people who will be there for 2 months or less. At
sample units where all the people are staying less than 2 months, the
respondent is asked a subset of the questions from the housing section,
including the question on vacancy status. The current residence concept
is unique to the American Community Survey. By using the current
residence to decide for whom to collect survey information, the ACS can
provide a more accurate description of an area's social and economic
characteristics. Most surveys, as well as the decennial census, use the
concept of usual residence. Usual residence is defined as the place
where a person lives and sleeps most of the time. The census defines
everyone as having only one usual residence.

## Methodology for estimating seasonal populations using ACS data

In order to reduce the complexity of estimates, this methodology
proposes, as an initial step, one particular group, snowbirds, seasonal
residents.

In order to develop a strategy for estimating seasonal populations
within CCDPH, we first carried out a literature review. American
Community Survey to derive estimates of the seasonal populations.

## Seasonal population estimates for CCDPH

Figure 1. Population pyramids, 2000 and 2019 Changes in the CCDPH
jurisdiction's age composition

```{r Population Pyramids, figures-side, fig.show="hold", out.width="50%"}

par(mar = c(4, 4, .1, .1))

fig01_pyramid_2010 %>% 
  mutate(Age= factor(Age,levels = c("PopUnd5",
                          "Pop5to9",
                          "Pop10to14",
                          "Pop15to19",
                          "Pop20to24",
                          "Pop25to29",
                          "Pop30to34",
                          "Pop35to39",
                          "Pop40to44",
                          "Pop45to49",
                          "Pop50to54",
                          "Pop55to59",
                          "Pop60to64",
                          "Pop65to69",
                          "Pop70to74",
                          "Pop75to79",
                          "Pop80to84",
                          "Pop85Over"))) %>%
  ggplot(aes(x = Age, y = Population, fill = sex)) +
  ggtitle("CCDPH Age Composition, 2010") +
  geom_bar(stat = "identity", width = 0.85) + 
  scale_y_continuous(limits= max(fig01_pyramid_2010$Population) * c(-1,1)) + 
  scale_fill_manual(values=as.vector(c("grey45","grey63"))) +
  coord_flip() +
  labs(x="",y="",fill="") + 
  theme(plot.title = element_text(hjust = 0.5))

fig01_pyramid_2019 %>% 
  mutate(Age= factor(Age,levels = c("PopUnd5",
                          "Pop5to9",
                          "Pop10to14",
                          "Pop15to19",
                          "Pop20to24",
                          "Pop25to29",
                          "Pop30to34",
                          "Pop35to39",
                          "Pop40to44",
                          "Pop45to49",
                          "Pop50to54",
                          "Pop55to59",
                          "Pop60to64",
                          "Pop65to69",
                          "Pop70to74",
                          "Pop75to79",
                          "Pop80to84",
                          "Pop85Over"))) %>%
  ggplot(aes(x = Age, y = Population, fill = sex)) +
  ggtitle("CCDPH Age Composition, 2019") +
  geom_bar(stat = "identity", width = 0.85) + 
  scale_y_continuous(limits= max(fig01_pyramid_2010$Population) * c(-1,1)) + 
  scale_fill_manual(values=as.vector(c("grey45","grey63"))) +
  coord_flip() +
  labs(x="",y="",fill="") + 
  theme(plot.title = element_text(hjust = 0.5))

```

Figure 2. Population 65 and older population in CCDPH, 2000-2019 (Table
B01001 Population by Age, Sex) Trends in the population 65 and older
population over time

Figure 3. Mobility status of 65 and older population in CCDPH, 2000-2019
(Table B01001 Population by Age, Sex) Trends in the population 65 and
older population over time

Figure 4. Seasonally vacant housing, 2000-2019 (Table B25005 VACANT -
CURRENT RESIDENCE ELSEWHERE) (Table B25004 VACANCY STATUS)

## Discussion

## References
